The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-04 rnw/Analysis and Visualization.Rnw
update Andreas 04.10.2017(2)

\begin{small}
<<>>=
extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                        small_area = list(sa.col = "smallarea", 
                                          areas = c("A", "B","C", "D"),
                                          unbiased = TRUE))
@
\end{small}

@@

\begin{small}
<<>>=
extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
                                unbiased = TRUE),
                boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
psynth_3p<- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area = list(sa.col = "smallarea", 
                                  areas = c("A", "B", "C", "D"),
                                  unbiased = FALSE),
                boundary_weights = "boundary_weights")
@
\end{small}

@@

\begin{small}
<<>>=
grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                             extpsynth_3p, psynth_3p),
                             sae = TRUE, 
                             vartypes = c("variance", "g_variance",  "ext_variance"))
@
\end{small}

>>>>

\begin{small}
<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                        small_area = list(sa.col = "smallarea", 
                                          areas = c("A", "B","C", "D"),
                                          unbiased = TRUE))
@
\end{small}

@@

\begin{small}
<<>>=
extpsynth_3p <- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
                                unbiased = TRUE),
                boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
psynth_3p <- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area = list(sa.col = "smallarea", 
                                  areas = c("A", "B", "C", "D"),
                                  unbiased = FALSE),
                boundary_weights = "boundary_weights")
@
\end{small}

@@

\begin{small}
<<>>=
grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                             extpsynth_3p, psynth_3p),
                             sae = TRUE, 
                             vartypes = c("variance", "g_variance",  "ext_variance"))
@
\end{small}

### 2017-10-04 rnw/Special Cases and Scenarios.Rnw
update Andreas 04.10.2017(2)

\begin{small}
<<>>=
grisons$stage<- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
@
\end{small}

@@

In oder to keep track of such cases, the package \pkg{forestinventory} tells the user to do so by returning a warning message:
<<echo=FALSE>>=
extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      cluster = "cluster",
                      small_area = list(sa.col = "ismallold", areas = c("1"),
                                        unbiased = TRUE))
@
\begin{small}
\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      cluster = "cluster",
+                      small_area = list(sa.col = "ismallold", areas = c("1"),
+                                        unbiased = TRUE))
Warning message:
At least one terrestrial cluster not entirely included within the small area 1.
Zero mean residual assumption for small area maybe violated.
Check mean_Rc_x_hat_G and consider alternative estimator 'psmall' 
\end{Sinput}
\end{Schunk}
\end{small}

\begin{small}
<<>>=
extpsmall.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      cluster = "cluster",
                      small_area = list(sa.col = "ismallold", areas = c("1"),
                                        unbiased = TRUE),
                     psmall = TRUE)
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "mean" value from an s2- (i.e. s1- and s0-) sample point:
grisons[which(grisons$phase_id_3p==2)[1],"mean"]<- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "q75" value from an s2-sample point:
grisons[which(grisons$phase_id_3p==2)[1],"q75"]<- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "mean" value from an s0-sample point:
grisons[which(grisons$phase_id_3p==0)[1],"mean"]<- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete s2-points with "stade"-level '300'
zberg<- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
@
\end{small}

>>>>

\begin{small}
<<>>=
grisons$stage <- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
@
\end{small}

@@

In oder to keep track of such cases, the package \pkg{forestinventory} tells the user to do so by returning a warning message:
<<echo=FALSE>>=
extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      cluster = "cluster",
                      small_area = list(sa.col = "ismallold", areas = c("1"),
                                        unbiased = TRUE))
@
\begin{small}
\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      cluster = "cluster",
+                      small_area = list(sa.col = "ismallold", areas = c("1"),
+                                        unbiased = TRUE))
Warning message:
At least one terrestrial cluster not entirely included within the small area 1.
Zero mean residual assumption for small area maybe violated.
Check mean_Rc_x_hat_G and consider alternative estimator 'psmall' 
\end{Sinput}
\end{Schunk}
\end{small}

\begin{small}
<<>>=
extpsmall.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      cluster = "cluster",
                      small_area = list(sa.col = "ismallold", areas = c("1"),
                                        unbiased = TRUE),
                     psmall = TRUE)
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "mean" value from an s2- (i.e. s1- and s0-) sample point:
grisons[which(grisons$phase_id_3p==2)[1],"mean"] <- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "q75" value from an s2-sample point:
grisons[which(grisons$phase_id_3p==2)[1],"q75"] <- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete "mean" value from an s0-sample point:
grisons[which(grisons$phase_id_3p==0)[1],"mean"] <- NA
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
## delete s2-points with "stade"-level '300'
zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
@
\end{small}

### 2017-10-04 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 04.10.2017(2)

\begin{small}
<<>>=
reg3p_nex<- threephase(formula.s0 = formula.rm, 
                  formula.s1 = formula.fm, data = grisons,
                  phase_id = list(phase.col="phase_id_3p", s1.id = 1, 
                                  terrgrid.id = 2),
                  boundary_weights = "boundary_weights")
@
\end{small}

>>>>

\begin{small}
<<>>=
reg3p_nex <- threephase(formula.s0 = formula.rm, 
                  formula.s1 = formula.fm, data = grisons,
                  phase_id = list(phase.col="phase_id_3p", s1.id = 1, 
                                  terrgrid.id = 2),
                  boundary_weights = "boundary_weights")
@
\end{small}
